Reproducible Research, Project 1
=============================

##### General / loading dependencies etc

```{r message=FALSE}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(gridExtra)
options(scipen=999) # displaying integers without scientific notation
```

##### Question 1, reading data
```{r}
dat <- read.csv("activity.csv",head=T)
# dat$date <- as.POSIXct(dat$date,"%Y-%m-%d",tz="GMT")
```


##### Question 2, histogram of the total number of steps take each day
```{r}
steps <-  dat %>% group_by(date) %>% summarise(steps.per.date=sum(steps,na.rm=T)) 
ggplot(data=steps, aes(steps.per.date))+
      geom_histogram(binwidth = 600)+
      labs(x="Total number of steps per day",y="Frequency ")+
      ggtitle("Histogram of the total number of steps taken per day")
```

##### Question 3, mean and median of steps taken each day
```{r, echo=T, results='hide'}
mean(steps$steps.per.date)
median(steps$steps.per.date)
```

Mean of steps taken each day is `r floor(mean(steps$steps.per.date))` and the Median is `r round(median(steps$steps.per.date))` 

##### Question 4, time series plot of the average number of steps taken
```{r}
avg.steps.per.interval <- 
      dat %>% 
      group_by(interval) %>%
      summarise(average.steps=mean(steps,na.rm=T))

ggplot(data=avg.steps.per.interval,aes(interval,average.steps))+
      geom_line()



```

##### Question 5, the 5-minute interval that, on average, contains the maximum number of steps


```{r}
max.interval <- subset(avg.steps.per.interval,avg.steps.per.interval$average.steps==max(avg.steps.per.interval$average.steps))
```

So interval `r max.interval$interval` is the period with the highest activity. It corresponds well with the peak on the graph from question 4. The average number of steps in that period was `r round(max.interval$average.steps)` steps


##### Question 6 and 7, code to describe and show a strategy for imputing missing data and Histogram

In the data set there are `r sum(is.na(dat))` missing values which corresponds to around 4.37% of the entire data set.

Filling the missing values with the average of each interval would result in an unchanged mean of the entire data set.

Again using the dplyr package:

```{r}
imp <- inner_join(dat, avg.steps.per.interval, by="interval") %>%
      mutate(steps=ifelse(is.na(steps),average.steps,steps)) %>%
      select(date,interval, steps)

dat.imp <- 
      imp %>%
      group_by(date) %>%
      summarise(steps.per.day.imp = sum(steps))
ggplot(data=dat.imp,aes(steps.per.day.imp))+
      geom_histogram(binwidth = 600)+
      labs(x="Steps per day",y="Frequency")+
      ggtitle("Histogram as in question 2 (with imputed data)")
```


Calculating mean and median after imputation.
```{r}
imp.dat2 <- 
      imp %>% 
      group_by(date) %>%
      summarise(steps.per.date=sum(steps,na.rm=T))
mean(imp.dat2$steps.per.date)
median(imp.dat2$steps.per.date)
```
Shows that mean `r round(mean(imp.dat2$steps.per.date))` is, as expected, the same as before but the median is now `r round(median(imp.dat2$steps.per.date))` higher than before (and equal to the mean). Apparently the process of imputing the missing values did not have a major impact on the mean / median of the data set. 

##### Question 8, panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Using the imputed data to look at the differences between weekday activity and weekend activity.
```{r}
imp$date <- as.POSIXct(imp$date)
imp$day <- weekdays(imp$date)
day.interval <-  
      imp %>%
      group_by(interval,day) %>%
      summarise(mean.w.day=mean(steps,na.rm=T))

```

Dividing days into weekdays and weekend-days

```{r}
weekday <- filter(day.interval, day=="Monday" | day== "Tuesday" | day == "Wednesday" | day == "Thursday" | day == "Friday")
weekendday <- filter(day.interval, day=="Saturday" | day == "Sunday")
```

```{r}
weekday.dat <- 
      weekday %>% 
      group_by(interval) %>%
      summarise(var1=mean(mean.w.day))
weekday.gg <- ggplot(data=weekday.dat, aes(interval, var1))+geom_line()+ylim(c(0,250))+ylab("Average steps")+ggtitle("Weekdays")

weekendday.dat <- 
      weekendday %>%
      group_by(interval) %>%
      summarise(var2=mean(mean.w.day))
weekendday.gg <- ggplot(data=weekendday.dat,aes(interval,var2))+geom_line()+ylim(c(0,250))+ylab("Average steps")+ggtitle("Weekend")

grid.arrange(weekday.gg,weekendday.gg,nrow=2)

```
